---
title: "Project 1"
author: "William Davis"
date: "6/11/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(caret)
library(xgboost)
library(doSNOW)
library(glmnet)
library(nnet)

```

## **Background**
### - In America, 1 in every 4 males suffer from heart disease related deaths each year.
### - This attributed to 357,761 deaths in 2019. I chose the Heart Attack Analysis and Prediction Dataset to help scientist and medical professional prevent future events of heart attack. 
### - This dataset will predict the likelihood of a paitient going into cardiac arrest. 

```{r heart data}
heart_data = read.csv("heart.csv", sep = ",")
heart_data$cp <- as.factor(heart_data$cp)
heart_data$sex <- as.factor(heart_data$sex)
heart_data$age <- as.factor(heart_data$age)
heart_data$restecg <- as.factor(heart_data$restecg)
heart_data$exng <- as.factor(heart_data$exng)
heart_data$fbs  <- as.factor(heart_data$fbs)
heart_data$output <- as.factor(heart_data$output)
heart_data$thall <- as.factor(heart_data$thall)

```

## The heart attack analysis and prediction dataset, used many fields that needed to cleaned and converted to factors. In this dataset the following values were converted into factors.
### - age
### - sex
### - cp (chest pain)
### - restecg (resting electrocardiographic results)
### - exng (exercise induced angina)
### - fbs (fasting blood sugar > 120 mg/dl)
### - thall 
### - output (likely-hood of having a heart attack)

# Feature Engineering
## To validate the data values for thallach (maximum heart rate acheived) and trtbps (resting blood pressure), The values have been normalized and used in the predictive model. 

```{r mean values}


meanMaxHeartRateAtt = mean(heart_data$thalachh)
s = sd(heart_data$thalachh)
heart_data$MaxHeartNormal = (heart_data$thalachh - meanMaxHeartRateAtt)/s


meanRestBlPress = mean(heart_data$trtbps)
s = sd(heart_data$trtbps)
heart_data$RestBloodPressNormal = (heart_data$trtbps - meanRestBlPress)/s

diabetes = heart_data$fbs

diabetesChance <- character()

for(i in 1:length(diabetes)){
  if(diabetes[i] == 1){
    diabetesChance[i] <- 1L
  }else if(diabetes[i] == 0){
    diabetesChance[i] <- 0L
  }
}


heart_data$twiceAsLikely = diabetesChance

```

## According to the CDC, men with diabetes are twice as likley to enter cardiac arrest. With this in mind, a paitent with a fbs above 120 is likely to have diabetes. The feature "twiceAsLikely" has been created and added to the dataset to contribute to the predictive model with 1 = twice as likely, and 2 = not twice as likely. 
  

## Data Sampling and Predictive Modeling

```{r Data Sampling Training Cluster}

heart_dataSelect <- select(heart_data,age, sex, caa, cp,chol, 
                           trtbps, fbs,thall,thalachh, exng, 
                           output, MaxHeartNormal, 
                           RestBloodPressNormal, twiceAsLikely)

samples<- createDataPartition(heart_data$output,
                            times = 1,
                            p = 0.80,
                            list = F)

heart_dataSelectTrain <- heart_dataSelect[samples,]
heart_dataSelectTest <- heart_dataSelect[-samples,]

control <- trainControl(method ="repeatedcv",
                        number = 10,
                        repeats = 2,
                        search = "grid")

cl <- makeCluster(3, type = "SOCK")
registerDoSNOW(cl)
set.seed(1234)
caret.cv <- train(output ~.,
                  data = heart_dataSelectTrain,
                  method = "multinom",
                  preProcess = ("center"),
                  trControl = control)

stopCluster(cl)



pred <- predict(caret.cv, heart_dataSelectTest)


```
## Multinominal Logistic Regression was used to determine the likelihood of a patient entering cardiac arrests. The predictive model was created using the caret and doSnow libraries. Multiple clusters were created to reduced the time to compute the results. 


# Results

## A confusion matrix was created to verify the accuracy of the data.

```{r Results Graphing}
confusionMatrix(pred, heart_dataSelectTest$output)

ggplot(heart_dataSelectTest, 
       aes(x = trtbps, 
           y  = thalachh, 
           col = pred)) + 
  geom_point() + 
  geom_smooth(method = lm) + 
  ggtitle("Chance of Heart Attack ") +
  xlab("Resting Heart Rate") + 
  ylab("Maximum Heart Rate")


```

## Plotting the resting heart rate vs the max rate, we can see that we predicted that if the as the maximum heart rate increases the chance of a patient entering cardiac arrest increase, even if the resting heart rate is lower. 
options(knitr.duplicate.label = "allow")
```{r Graphing Heart}
ggplot(heart_dataSelectTest, 
       aes(x = trtbps, 
           y  = thalachh, 
           col = output)) + 
  geom_point() + 
  geom_smooth(method = lm) + 
  ggtitle(" Chance of Heart Attack ") +
  xlab("Resting Heart Rate") + 
  ylab("Maximum Heart Rate")


```
## The same is true for the initital un-predicted outcome.

```{r Graphing Pred}
ggplot(heart_dataSelectTest, 
       aes(x = trtbps, 
           y  =thalachh, 
           col = twiceAsLikely)) + 
  geom_point() + 
  geom_smooth(method = lm) + 
  ggtitle("Chance of Heart Attack ") +
  xlab("Resting Heart Rate") + 
  ylab("Maximum Heart Rate")

```
#Dataset

### https://www.kaggle.com/datasets/rashikrahmanpritom/heart-attack-analysis-prediction-dataset/discussion/234843

#References

### D. (2020, May 28). Multinomial logistic regression With R. R-Bloggers. https://www.r-bloggers.com/2020/05/multinomial-logistic-regression-with-r/#:%7E:text=Overview%20%E2%80%93%20Multinomial%20logistic%20Regression,to%20use%20ordinal%20logistic%20regression.Dalpiaz, D. (2020, October 28). Chapter 21 The caret Package | R for Statistical Learning. Github.Io.

### https://daviddalpiaz.github.io/r4sl/the-caret-package.htmlDiabetes and Your Heart. (2021, May 7). Centers for Disease Control and Prevention.

### https://www.cdc.gov/diabetes/library/features/diabetes-and-heart.html#:%7E:text=If%20you%20have%20diabetes%2C%20you,are%20to%20have%20heart%20disease.ECG test - Better Health Channel. (2022). ECG Test - Better Health Channel. Retrieved June 13, 2022, from https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/ecg-testHeart Attack Analysis & Prediction Dataset. (2021, March 22). Kaggle.

### https://www.kaggle.com/datasets/rashikrahmanpritom/heart-attack-analysis-prediction-dataset/discussion/234843Men and Heart Disease. (2022, March 8). Centers for Disease Control and Prevention.

### https://www.cdc.gov/heartdisease/men.htm#:%7E:text=About%201%20in%2013%20(7.7,men%20have%20coronary%20heart%20disease.&text=Half%20of%20the%20men%20who,disease%20had%20no%20previous%20symptoms.&text=Even%20if%20you%20have%20no,at%20risk%20for%20heart%20disease.Prediabetes - Diagnosis and treatment - Mayo Clinic. (2022, March 17).

### Mayoclinic.Org. https://www.mayoclinic.org/diseases-conditions/prediabetes/diagnosis-treatment/drc-20355284#:%7E:text=Fasting%20blood%20sugar%20test&text=L)%20of%20blood.-,In%20general%3A,tests%20is%20diagnosed%20as%20diabetes


